The supplementary material of ``Computing solution space properties of combinatorial optimization problems via generic tensor networks''

### 2022-06-06 paper/suppl.tex
update

\documentclass[onefignum, onetabnum]{siamart190516}
%\usepackage{natbib}
%\usepackage[sort]{cite}
\pdfoutput=1
%\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{quoting}
\usepackage{upquote}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[framemethod=TikZ]{mdframed}
\usetikzlibrary{shapes}
\usetikzlibrary{snakes}
\usetikzlibrary{shapes.geometric}
\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage[plain]{algorithmic}
\usepackage[linesnumbered, ruled, vlined, algo2e]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{rotating}
%\usepackage{cite}
\usepackage{booktabs}
%\usepackage{unicode-math}
%\usepackage{algorithm}% http://ctan.org/pkg/algorithm
%\usepackage{algpseudocode}% http://ctan.org/pkg/algpseudocode
\usepackage{xcolor}% http://ctan.org/pkg/xcolor
\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother
\usepackage{bbm}
\usepackage{jlcode}
\usepackage{graphicx}
\usepackage{amsmath,color}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{indentfirst}
\usepackage{txfonts}
%\usepackage[epsilon, tsrm, altpo]{backnaur}

@@

\title{Computing solution space properties of combinatorial optimization problems via generic tensor networks
%\thanks{\funding{...}}
}

\author{
Jin-Guo Liu\thanks{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA; QuEra Computing Inc., 1284 Soldiers Field Road, Boston, MA, 02135, USA  %\thanks{QuEra Computing Inc., 1284 Soldiers Field Road, Boston, MA, 02135, USA}
    (\email{jinguoliu@g.harvard.edu}).}
\and {Xun Gao}\thanks{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA (\email{xungao@g.harvard.edu}, contributed equally with Jin-Guo Liu to this work.).}
\and Madelyn Cain\footnotemark[2]
\and Mikhail D. Lukin\footnotemark[2]
\and Sheng-Tao Wang\thanks{QuEra Computing Inc., 1284 Soldiers Field Road, Boston, MA, 02135, USA}
}

@@

\begin{abstract}
We introduce a unified framework to compute the solution space properties of a broad class of combinatorial optimization problems.
These properties include finding one of the optimum solutions, counting the number of solutions of a given size, and enumeration and sampling of solutions of a given size. Using the independent set problem as an example, we show how all these solution space properties can be computed in the unified approach of generic tensor networks.
We demonstrate the versatility of this computational tool by applying it to several  examples, including computing the entropy constant for hardcore lattice gases, studying the overlap gap properties, and analyzing the performance of quantum and classical algorithms for finding maximum independent sets.
\end{abstract}

% REQUIRED
\begin{keywords}
solution space property, tensor networks, maximum independent set, independence polynomial, generic programming, combinatorial optimization 
\end{keywords}

% REQUIRED
% 14N07  	Secant varieties, tensor rank, varieties of sums of powers
\begin{AMS}
  15A69, 05C31, 14N10
\end{AMS}

\section{Introduction}

An important class of problems in graph theory and combinatorial optimization  can be formulated as satisfiability problems involving constraints specified over a vertex and its neighborhood.
These include, for example, the independent set, the cutting problem, the dominating set, the set packing and  the set covering, the vertex coloring problem, the K-SAT,  the clique problem, and the vertex cover problem~\cite{Moore2011}.
These problems have a wide range of applications in scheduling, logistics,
wireless networks and telecommunication, and computer vision, among others~\cite{Butenko2003, Wu2015}.
Finding an optimum solution for these problems is typically NP-hard~\cite{Hastad1996}.
%, which implies they can not to be solved in a time that scales polynomialy  with the size of the problem.

% The most studied aspect of these problems is to find an optimum solution, 
%The most studied version of the above problems is finding an optimal solution maximum/minimum set satisfying the contraints and the corresponding set size,
% which is typically 
%while many of them belong to the hardness category
% NP-hard~\cite{Hastad1996}, i.e.\ it is unlikely to be solved in a time polynomial with the size of the problem.

% Many properties of these problems are in general intractable at large sizes (exponential run time in problem size in the worst case),
 %but it is nevertheless desirable to have high-performance algorithms because of their ubiquitous applications~\cite{Butenko2003, Wu2015}.

%\purple{Is MIS really the most studied NP-hard optimization problem? I would guess it is heavily studied, but not necessarily the most studied.}
%\blue{Here, we haven't mentioned the independent set, here ``version'' means, e.g. counting, enumeration, devision. I think finding one of the largest/smallest set and its cardinality in the above mentioned problems is the most studied?}

In this Article, we introduce a unified framework to compute a broad class of properties 
associated with the solutions of these problems, beyond just finding an optimum solution.
We call them \textit{solution space properties}. In practice, these can be much harder to compute (corresponding e.g.\ to \#P-complete class~\cite{Moore2011}). However, these properties can be crucial for understanding detailed properties of hard combinatorial optimization problems.
% In this paper, we focus on the less studied aspects of computing \textit{solution space properties}, which can be much harder to compute, %; for example, counting independent sets of a general graph is \#P-complete~\cite{Vadhan2001}, 
% but they can be crucial towards understanding hard combinatorial optimization problems. 
These \textit{solution space properties} can include not only the maximum or minimum set size but also the number of sets at a given size, enumeration of all sets at a given size, and direct sampling of such sets when they are too large to be fit into memory.
% For the weighted version of the above problems, it also includes finding the largest or smallest $k$ sets and their sizes.
%are many interesting and hard computational problems concerning various properties of independent sets.
%Solution space property computation allows us to understand the solution space better.
They can be used to understand the hardness of finding an optimum solution for a given problem instance and the performance of a specific solver.  
%in computing the solution for the given instance.
For example, the number of configurations at different sizes can inform how likely a simulated annealing algorithm will be trapped in local minima at certain sizes~\cite{Xu2018}.
%In the past, people use the counting of configurations at different sizes to tell how likely the simulated annealing will be trapped in certain size in the strong defensive alliance problem~\cite{Xu2018}, 
The pair-wise Hamming distance distribution of configurations at a given size can indicate the presence or absence of the overlap gap property~\cite{Gamarnik2013, Gamarnik2019}, which can be used to bound the performance of local optimization algorithms.
%a no-go theorem about when the solution space geometry is formidably
%hard for any local search algorithm.
In a recent experiment based on a Rydberg atom array quantum computer, 
%the authors use 
the counting and the configuration space connectivity information was used to find maximum independent set (MIS) problem instances that are hard for simulated annealing and to evaluate the corresponding quantum algorithm performance \cite{Ebadi2022}.
% In a recent experiment using a Rydberg atom array quantum simulator to find maximum independent sets (MIS) of a graph~\cite{Ebadi2022}, the authors use the counting information and the configuration space connectivity to find instances hard for simulated annealing. %and design better quantum algorithms 
%to solve these instances faster, 
The need for understanding these important aspects of combinatorial optimization motivates us to find methodologies to compute these solution space properties.
%\purple{I would give examples of why we want to compute solution space properties.
%For instance, in statistical physics, computing thermal states and
%observables via the partition function. In complexity theory, proving limits on local algorithms (OGP).}\blue{now?}

To this end, we show how to obtain all of these seemingly unrelated properties in a unified approach using
\textit{generic tensor networks}. Tensor networks are a computational model widely used in condensed matter physics~\cite{Orus2014}, quantum computing~\cite{Markov2008}, big data~\cite{Cichocki2014} and mathematics~\cite{Oseledets2011}.
They are also known as the sum-product networks in probabilistic modeling~\cite{Bishop2006} or \texttt{einsum} in linear algebra libraries such as NumPy~\cite{Harris2020}.
%\purple{I’d write out $2^{{\rm tw}(G)}$}\blue{I feel it is hard to explain line graph without texts.}
Recent progress in simulating quantum circuits with tensor networks~\cite{Gray2021, Pan2021, Kalachev2021} makes it possible to contract a randomly structured sparse tensor network with up to thousands of tensors in a reasonable time.
%\purple{Is the scaling of these algorithms known?}\blue{They are not exact algorithms, the time can be any, typically a few minutes, will it be helpful to mention this?}
In previous studies, the data types of the tensor elements are typically restricted to standard number types such as real numbers and complex numbers.
Here, we extend to \textit{generic tensor networks} by generalizing the tensor element data types to any type that has the algebraic structure of a commutative semiring.
%These two algorithms are related but are not equivalent.
%Dynamic programming is more general purpose, while a tensor network has a richer algebraic structure, and this structure is crucial for solution space property computation.
%\purple{Not clear how first part and second part of sentence are related}\blue{now? or maybe just removing the comparison with dynamic programming?}
%For sparse graphs, the treewidth is usually much smaller than the number of vertices~\cite{Fomin2006}. Our algorithms, on the other hand, are much more versatile than traditional methods and can be used to compute many other properties than just the MIS size.
%The independent set problem is related to physics applications such as in the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007}
%in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018, Ebadi2022};
%Property computation can, for example, be used to understand phase transitions~\cite{Dyre2016, Fernandes2007},
%to identify harder graphs in an ensemble of graphs~\cite{Ebadi2022} and to analyze the presence of the overlap gap property~\cite{Gamarnik2013, Gamarnik2019} related to the clustering phenomenon of large independent sets.
%However, there is a lack of general and versatile tool to compute these different properties of independent sets.
%To keep the discussion focused, 
In what follows, for clarity of presentation, we focus on the independent set problem in the main text, and show how to compute the solution space properties for other combinatorial optimization problems in  \App{sec:otherproblems}. The latter include cutting, matching, vertex coloring, satisfiability,  dominating set,  set packing,  set covering, and the clique problem.
%\xpurple{More details on which combinatorial optimization problems this applies to?}

The paper is organized as follows.
We first introduce the basic concepts of tensor networks and generic programming in \Sec{sec:tn} and \Sec{sec:generic}.
Then we show how to reduce the independent set problem to a tensor network contraction problem in \Sec{sec:tnmap}. 
Subsequently, we explain how to engineer the element types to compute various solution space properties in \Sec{sec:counting}, \Sec{sec:enumeration}, and \Sec{sec:weighted}. % to count independent sets at a given size, enumerate/sample independent sets at a given size and find largest set sizes in weighted graphs, respectively.
% by computing a number of independent set properties of certain sparse graphs on central processing units (CPUs) and graphics processing units (GPUs) and show its good performance.
Lastly, we benchmark our algorithms in \Sec{sec:benchmark} and provide three example applications in \Sec{sec:examples} to demonstrate the versatility of our tool.
%In the first example, we show how our method can help computing the entropy constant for some hardcore lattice gases on 2D square lattices. We compute the 2D generalization of Fibonacci integer sequence up to size $39$, which is larger than the previous record $37$.
%In the second example, we analyze the pairwise Hamming distances from configurations directly sampled from large independent sets of King's graphs at $0.8$ filling and three regular graphs. We show the strongest numeric evidence so far of the absence of overlap gap property in King's graph and the existence of Overlap Gap properties in three regular graphs.
%In the last example, we show how our algorithm helps people to understand the performance of a quantum variational algorithms on Rydberg atom arrays to find maximum independent sets. This is a research frontier that understanding which can help people design better quantum algorithms to beat its classical counterpart.
%Our method can also be used to find ``maximal'' independent set properties; 
%a maximal independent set is an independent set that is not a subset of any other independent set, but its size may not be the maximum.

\section{Tensor networks}\label{sec:tn}
A tensor network~\cite{Cirac2020, Orus2014} is composed of a collection of tensors, a collection of labels associated with tensor dimensions, and an indicator of which tensor is the output.
The operation to evaluate a tensor network is called \textit{contraction}, which is defined as a summation of tensor element products over the labels not appearing in the output tensor.
For example, the matrix multiplication is a special tensor network that can be represented as $C_{ik} = A_{ij}B_{jk}$, where $A, B$ and $C$ are matrices (two-dimensional tensors), $ik,ij$ and $ik$ are labels associated with them, and ``$=$'' is the indicator of $C$ being the output, while its contraction is defined as $C_{ik} = \sum_j A_{ij}B_{jk}$.
%can be viewed as a generalization of binary matrix multiplication to nary tensor contraction.
%Generalized Einstein's notation is often used to represent a tensor network, where we use a label in the subscript to represent a degree of freedom.
%One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% Einstein's notation is often used to represent a tensor network, e.g.\ it represents the matrix multiplication between two matrices $A$ and $B$ as $C_{ik} = A_{ij}B_{jk}$,
% where we use a label in the subscript to represent a degree of freedom.
% One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
% In the standard Einstein's notation for tensor networks in physics, each index appears precisely twice, either both in input tensors (which will be summed over) or one in a input tensor and another in the output tensor.
% Hence a tensor network can be represented as a open simple graph,
% where a tensor is mapped to a vertex, a label shared by two input tensors is mapped to a normal edge and a label appears in the output tensor is mapped to an open edge.
% In this work, we adopt a more general form of tensor networks that do not restrict how many times a label can appear in tensors. 
The graphical representation of a tensor network is an open (meaning an edge can connect to external vertices) hypergraph, where an input tensor is mapped to a vertex and a label is mapped to a hyperedge that can connect an arbitrary number of vertices, while the labels appearing in the output tensor are open.
Our notation is a minor generalization of the standard tensor network notation used in physics as we do not restrict the number of times a label can appear in the tensors. 
While this generalized form is equivalent in representation power, it can have smaller contraction complexity as will be illustrated in \App{sec:tensorbad}.
% because one can easily translate a generalized tensor network to the standard notation by adding $\delta$ tensors, a high dimensional equivalence of the identity matrix.
%However, introducing $\delta$ tensors may significantly increase the complexity of the tensor contraction. We illustrate this subtle point in \App{sec:tensorbad}.

\begin{example}
$C_{ijk} = A_{jkm} B_{mil} V_{jm}$ is a tensor network that can be evaluated as $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$.
Its hypergraph representation is shown below, where we use different colors to represent different hyperedges.

\vspace{1em}
\centerline{\begin{tikzpicture}[
    dot/.style = {circle, fill, minimum size=#1,
                inner sep=0pt, outer sep=0pt},
    dot/.default = 6pt  % size of the circle diameter 
                    ]  
    \def\dx{0};
    \def\r{0.5cm}
    \def\sr{0.15cm}
    \def\ax{0}
    \def\ay{0}
    \def\bx{1}
    \def\by{1}
    \def\cx{1}
    \def\cy{-1}
    \node[color=white,fill=black,dot=\r] at (\ax+\dx,\ax) (a) {A};
    \node[color=white,fill=black,dot=\r] at (\bx+\dx,\by) (b) {B};
    \node[color=white,fill=black,dot=\r] at (\cx+\dx,\cy) (v) {V};
    \node[color=transparent,draw=transparent,dot=0] at (\ax-1,\by) (o1) {};
    \node[color=transparent,draw=transparent,dot=0] at (\ax-1.5,\ay) (o2) {};
    \node[color=transparent,draw=transparent,dot=0] at (\ax-1,\cy) (o3) {};
    \node at (\ax-0.8,\ay) (k) {k};
    \node at (\bx+0.4,\ay) (m) {m};
    \node at (\ax,\cy) (j) {j};
    \node at (\bx+1,\by) (l) {l};
    \node at (\bx-1,\by) (i) {i};
    \draw[color=blue,thick] (i) -- (b);
    \draw[color=blue,thick] (i) -- (o1);
    \draw[color=cyan,thick] (l) -- (b);
    \draw[color=violet,thick] (k) -- (a);
    \draw[color=violet,thick] (k) -- (o2);
    \draw[color=black,thick] (b) -- (m);
    \draw[color=black,thick] (m) -- (a);
    \draw[color=black,thick] (m) -- (v);
    \draw[color=red,thick] (a) -- (j);
    \draw[color=red,thick] (v) -- (j);
    \draw[color=red,thick] (o3) -- (j);
\end{tikzpicture}}
\end{example}

\section{Generic programming tensor contractions}\label{sec:generic}
%\purple{On a second read of this section, I think that the clarity could be improved. My understanding is this is the flow: We would like to use generic programming, where we write code that works regardless of input type, and is optimized on all input types without sacrificing efficiency. In general, we can write code such that a function runs correctly regardless of the input data type. This is easy to do for dynamically typed languages like Python, which only verify and enforce the rules for the input data type at run time (called type-checking). However, when the input data type is only checked at run time (as opposed to when the code is compiled), type-specific optimization cannot be used. High efficiency is only achieved in statically typed languages, where modern compiling technology allows the function to be individually optimized for each input data type (then give C++/Julia examples).}\blue{how about now?}
In previous works relating tensor networks and combinatoric problems~\cite{Kourtis2019, Biamonte2017}, the element types in the tensor networks are limited to standard number types such as floating-point numbers and integers.
%Owning to the development of modern compiling technology, we no longer need to restrict our imagination to standard number types.
%However, the elements in a tensor do not have to be a regular number type, or even a field (since division is not used in tensor contraction).
%One can use the same tensor contraction program for different purposes with different tensor element types.
% We propose to not restricting our imagination to standard number types, but to use a more general element type with certain algebraic property.
We propose to use more general element types with a certain algebraic property.
With different data types, we can solve different problems within the same unified framework.
This idea of using the same program for different purposes is also called generic programming in computer science:

\begin{definition}[Generic programming ~\cite{Stepanov2014}]
   Generic programming is an approach to programming that focuses on designing algorithms and data structures so that they work in the most general setting without loss of efficiency.
\end{definition}
This definition of generic programming covers two major aspects: ``work in the most general setting'' and ``without loss of efficiency''.
By the most general setting, we mean that a single program should work correctly for the most general input data types. 
%\purple{I think the following paragraph could be clarified. When you read the definition of generic programming, it's not clear what ``in the most general setting'' means. My understanding is it just means that the program will work regardless of the input data type - it would be good to say this explicitly. }
%\blue{This definition is from the book I cited, I also do not like ``in the most general setting'', maybe we just use our version?}
%\purple{I also think the way it is written takes a while to understand what dynamically versus statically typed languages are. The definition, “When these languages “see” a new input type, the compiler can recompile the generic program for the new type.” comes rather late in the paragraph.}
%\xpurple{Also, while I know what you mean, concepts like “inlining immutable elements” and “cache miss rate” would be
%lost on someone only with a physics background.}
%This definition of generic programming contains two major aspects: a single program works in the most general setting and efficiency.
%To understand the first aspect on generality,
For example, suppose we want to write a function that raises an element to a power, $f(x, n) := x^n$.
One can easily write a function for standard number types that computes the power of $x$ in $O \left( \log(n) \right)$ steps using the multiply and square trick.
Generic programming does not require $x$ to be a standard number type;
instead, it treats $x$ as an element with an associative multiplication operation $\odot$ and a multiplicative identity $\mymathbb{1}$.
Then, when the program takes a matrix as an input instead of a standard number type, it computes the matrix power correctly without rewriting the program.
The second aspect is about efficiency. For dynamically typed languages such as Python, the type information is not available for type-specific optimizations at the compilation stage.
Therefore, one can easily write code that works for general input types, but the efficiency is not guaranteed; for example, the speed of computing the matrix multiplication between two NumPy arrays with Python objects as elements is much slower than statically typed (i.e. the type information can be accessed at the compilation stage) languages such as C++ and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia takes advantage of just-in-time compilation and multiple dispatches.
When these languages ``see'' a new input type, the compiler recompiles the generic program for the new type to generate an efficient binary.
A myriad of optimizations can be done during the compilation. For example, the compiler can optimize the memory layout of immutable elements with fixed sizes in an array to speed up array indexing.
In Julia, if a type is immutable and contains no references to other values, an array of that type can even be compiled to graphics processing units (GPU) for faster computation~\cite{Besard2018}.

This motivates us to identify the most general tensor element type allowed in a tensor network contraction program.
We find that as long as the tensor elements are members of a commutative semiring, the tensor network contraction will be well defined and the result will be independent of the contraction order.
Here, a commutative semiring is a field that needs not to have an additive inverse and a multiplicative inverse.
Giving up these nice properties of fields has significant implications for tensor computation: tensor network compression algorithms might not be applicable because matrix factorization is NP-hard for commutative semirings~\cite{Shitov2014} and matrix multiplication faster than $O(n^3)$ does not exist for an algebra without an additive inverse~\cite{Kerr1970}.
Here, we only use the commutative properties of an algebra for optimizing the tensor network contraction order.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $S$, the following relations must hold for any arbitrary three elements $a, b, c \in S$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
%We require the algebra being multiplicative commutative because the contraction order optimizer will change the multiplication order of elements, while we require that the contraction order does not change the contraction results.
In the following sections, we will show how to compute solution space properties of independent sets using the same tensor network contraction algorithm by engineering tensor element algebra.
The Venn diagram in \Fig{fig:venn-diagram} shows the different types of algebra we will introduce in the main text and their relation, and Table~\ref{tbl:generictypes} summarizes which solution space properties can be computed by which tensor element types.

\begin{figure}[th]
   \centering
\centerline{\begin{tikzpicture}[]
    \scriptsize
    \node[draw,fill=lime!80,fill opacity=1, text opacity=1.0,ellipse,minimum width=2cm, minimum height=1cm,inner sep=0pt] at (0, 2.5) (R) {$\mathbbm{R}$};
    \def\dx{-3};
    \node[draw,fill=teal!50,fill opacity=1, text opacity=1.0,ellipse,minimum width=5cm, minimum height=3cm,inner sep=0pt] at (\dx, 0.15) (PN) {\hspace{-3.7cm}Polynomial};
    \node[draw,fill=brown!75,fill opacity=1, text opacity=1.0,ellipse,minimum width=3.5cm, minimum height=1.0cm,inner sep=0pt] at (\dx, 0.7) (P1) {\hspace{-1.0cm}Largest order};
    \node[draw,fill=brown!40,fill opacity=1, text opacity=1.0,ellipse,minimum width=3.5cm, minimum height=1.0cm,inner sep=0pt] at (\dx, -0.4) (P2) {\hspace{-0.8cm}Largest 2 orders};
    \node[draw,fill=brown,fill opacity=1, text opacity=1.0,ellipse,minimum width=0.8cm, minimum height=0.3cm,inner sep=0pt] at (\dx, 1.0) (T) {};
    \node[draw,fill=brown!50,fill opacity=1, text opacity=1.0,ellipse,minimum width=0.8cm, minimum height=0.3cm,inner sep=0pt] at (\dx, -0.1) (T2) {};
    \node at (\dx, -1.2) {$\ldots$};
    \node[above = 1cm] at (T) (textT) {Tropical};
    \node[below = 2cm, left=4cm] (textT2) {Extended Tropical};
    \draw[black,-latex] (textT) -- (T);
    \draw[black,-latex] (textT2) -- (T2);

    % set and set sampler
    \node[draw,fill=yellow,fill opacity=0.5, text opacity=1.0,ellipse,minimum width=5cm, minimum height=3cm,inner sep=0pt] at (0, 0) (SN) {};
    \node[below of=1] at (SN) {Set};
    \node[draw,fill=red!70,fill opacity=0.5, text opacity=1.0,ellipse,minimum width=4cm, minimum height=1.5cm,inner sep=0pt] at (-0.5, 0) (S1) {\hspace{1.5cm}Bit string};
\end{tikzpicture}}

    \caption{The tensor network element types used in this work and their relations.
    The overlap between two ellipses indicates that a new algebra can be created by combining those two types of algebra. ``Largest order'' and ``Largest 2 orders'' mean truncating the polynomial by only keeping its largest or largest two orders.
    The purposes of these element types can be found in \Tbl{tbl:generictypes}.}
     \label{fig:venn-diagram}
\end{figure}

\begin{table}[t!]\centering
\begin{minipage}{\columnwidth}
\ra{1.3}
        \begin{tabularx}{\textwidth}{sb}\toprule
            \hline
   \textbf{Element type}     & \textbf{Solution space property} \\
   {$\mathbbm{R}$}     & {Counting of all independent sets} \\
   {Polynomial} (\Eq{eq:polynomial})     & {Independence polynomial} \\
   {Tropical (\Eq{eq:tropical})}    & {Maximum independent set size} \\
   {Extended tropical of order $k$ (\Eq{eq:ext-tropical})}    & {Largest $k$ independent set sizes} \\
   {Polynomial truncated to $k$-th order (\Eq{eq:countingtropical} and \Eq{eq:max2poly})}     & {$k$ largest independent sizes and their degeneracy} \\
   {Set} (\Eq{eq:set})     & {Enumeration of independent sets} \\
   {Sum-Product expression tree} (\Eq{eq:expr})     & {Sampling of independent sets} \\
   {Polynomial truncated to largest order combined with bit string} (\Eq{eq:singleconfig})     & {Maximum independent set size and one of such configurations} \\
   {Polynomial truncated to $k$-th order combined with set} (\Eq{eq:countingtropicalset})    & {$k$ largest independent set sizes and their enumeration} \\
            \bottomrule
        \end{tabularx}
    \caption{Tensor element types and the independent set properties that can be computed using them.}\label{tbl:generictypes}
\end{minipage}
\end{table}

\section{Tensor network representation of independent sets} \label{sec:tnmap}
This section describes the reduction of the independent set problem to a tensor network contraction problem.
An alternative interpretation, perhaps more accessible to physicists, can be found in \App{sec:energymodel}, where we introduce the reduction from the energy model of hardcore lattice gases~\cite{Dyre2016, Fernandes2007}.
Let $G=(V,E)$ be an undirected graph with each vertex $v\in V$ being associated with a weight $w_v$.
An independent set $I \subseteq V$ is a set of vertices that for any vertex pair $u,v \in I$, we have $(u, v) \not\in E$.
To reduce the independent set problem on $G$ to a tensor network contraction, we first map a vertex $v\in V$ to a label $s_v \in \{0, 1\}$ of dimension $2$, where we use $0$ ($1$) to denote a vertex is absent (present) in the set.
For each vertex $v$, we define a parameterized rank-one tensor indexed by $s_v$ as
\begin{equation}
    W(x_v^{w_v}) = \left(\begin{matrix}
        1 \\
        x_v^{w_v}
    \end{matrix}\right),\label{eq:vertextensor}
\end{equation}
where $x_v$ is a variable associated with $v$.
Tensor elements can be indexed by subscripts, e.g.\ $W(x_v^{w_v})_0=1$ is the first element associated with $s_v=0$ and $W(x_v^{w_v})_1=x_v^{w_v}$ is the second element associated with $s_v=1$.
Similarly, for each edge $(u, v) \in E$, we define a matrix $B$ indexed by $s_u$ and $s_v$ as
\begin{equation}
    \qquad \quad 
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:edgetensor}
\end{equation}
which encodes the independent set constraint since $B_{11} = 0$, meaning the two vertices cannot be both in the independent set.

The corresponding tensor network contraction can be defined as
\begin{equation}\label{eq:idp}
    P(G, \{x_1^{w_1}, x_{2}^{w_2}, \ldots,x_{|V|}^{w_{|V|}}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{v\in V} W(x_v^{w_v})_{s_v} \prod\limits_{(u,v) \in E} B_{s_u s_v},
\end{equation}
where the summation runs over all $2^{|V|}$ vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$. 
% It is easy to see that an edge tensor element $B_{s_us_v}$ encodes the independent set constraint. It zeros out the product whenever both $u$ and $v$ are in the independent set.
A vertex tensor element $W(x_v^{w_v})_{s_v}$ contributes a multiplicative factor $x_v^{w_v}$ whenever $v$ is in the set.

\begin{example}\label{eg:twonode}
Here, we show a minimum example of mapping the independent problem of a 2-vertex complete graph K2 (left) to a tensor network (right).

\centerline{\begin{tikzpicture}[
dot/.style = {circle, fill, minimum size=#1,
            inner sep=0pt, outer sep=0pt},
dot/.default = 6pt  % size of the circle diameter 
                ]  
    \def\dx{0};
    \def\r{0.5cm}
    \def\wr{0.25cm}
    \node[dot=\r, fill=black] at (\dx,0) (a) {\color{white}{a}};
    \node[dot=\r, fill=black] at (2.5+\dx,0) (b) {\color{white}{b}};
    \draw [black,thick] (a) -- (b);

    \def\dx{5};
    \def\r{0.25cm}
    \foreach \x/\y/\e in {1.25/0/ab}
        \node[color=black,fill=black,dot=2*\r] at (\x+\dx,\y) (\e) {\scriptsize \color{white}{B}};
    \foreach \x/\y/\v in {0.3/0/a, 2.2/0/b}
        \node[color=black] at (\x+\dx,\y) (s\v) {$s_\v$};
    \foreach \x/\y/\v in {-0.5/0/a, 3.0/0.0/b}{
        \node[dot=\wr, color=black] at (\x+\dx,\y) (w\v) {};
        \node at (\x+\dx,\y+0.5) () {\scriptsize \color{black}{$W(x_\v^{w_\v})$}};
    }
    \draw [cyan,thick] (wa) -- (sa);
    \draw [cyan,thick] (sa) -- (ab);
    \draw [red,thick] (wb) -- (sb);
    \draw [red,thick] (sb) -- (ab);
\end{tikzpicture}}

In the graphical representation of the tensor network on the right panel,
we use a circle to represent a tensor, a hyperedge in cyan color to represent the degree of freedom $s_a$,
and a hyperedge in red color to represent the degree of freedom $s_b$.
Tensors sharing the same degree of freedom are connected by the same hyperedge.
The contraction of this tensor network has the following form:
\begin{equation}
    \begin{split}
    P({\rm K2},\{x_a^{w_a}, x_b^{w_b}\}) &=
    \left(\begin{matrix} 1 & x_a^{w_a} \end{matrix}\right)
    \left(\begin{matrix} 1 & 1\\ 1  & 0 \end{matrix}\right)
    \left(\begin{matrix} 1 \\ x_b^{w_b} \end{matrix}\right)\\
    &= 1 + x_a^{w_a} + x_b^{w_b}.
    \end{split}
\end{equation}
The resulting polynomial represents 3 different independent sets $\{\}$, $\{a\}$, and $\{b\}$, with weights being $0$, $w_a$, and $w_b$ respectively.
\end{example}

>>>>

%\documentclass[onefignum, onetabnum]{siamart190516}
\documentclass[a4paper, 12pt]{article}
%\usepackage{natbib}
%\usepackage[sort]{cite}
\pdfoutput=1
\usepackage[colorlinks=true,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{quoting}
\usepackage{upquote}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{cancel}
\usepackage[framemethod=TikZ]{mdframed}
\usetikzlibrary{shapes}
\usetikzlibrary{snakes}
\usetikzlibrary{shapes.geometric}
\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage{algorithmic}
\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{rotating}
%\usepackage{cite}
\usepackage{booktabs}
%\usepackage{unicode-math}
%\usepackage{algorithm}% http://ctan.org/pkg/algorithm
%\usepackage{algpseudocode}% http://ctan.org/pkg/algpseudocode
\usepackage{xcolor}% http://ctan.org/pkg/xcolor
\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother
\usepackage{bbm}
\usepackage{jlcode}
\usepackage{graphicx}
\usepackage{amsmath,color}
\usepackage{mathrsfs}
\usepackage{float}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{indentfirst}
\usepackage{txfonts}
\setcounter{secnumdepth}{0}
%\usepackage[epsilon, tsrm, altpo]{backnaur}

@@

\title{The supplementary material of ``Computing solution space properties of combinatorial optimization problems via generic tensor networks''
%\thanks{\funding{...}}
}

\author{
Jin-Guo Liu\thanks{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA; QuEra Computing Inc., 1284 Soldiers Field Road, Boston, MA, 02135, USA.}
\and {Xun Gao}\thanks{Department of Physics, Harvard University, Cambridge, Massachusetts 02138, USA (contributed equally with Jin-Guo Liu to this work.).}
\and Madelyn Cain\footnotemark[2]
\and Mikhail D. Lukin\footnotemark[2]
\and Sheng-Tao Wang\thanks{QuEra Computing Inc., 1284 Soldiers Field Road, Boston, MA, 02135, USA}
}

@@

\tableofcontents

% REQUIRED
% 14N07  	Secant varieties, tensor rank, varieties of sums of powers
